---
title: ""
message: false
warning: false
editor_options: 
  chunk_output_type: inline
---

# Logistic and Ensemble Models

## Libraries

Load Required Libraries

```{r}
library(tidyverse)
library(purrr)
library(gsheet)
library(raster)
library(ncdf4)
library(lubridate)
library(readxl)
library(writexl)
library(caret)
library(tidyr)
library(r4pde)
library(refund)
library(readr)
library(fdatest)
library(dplyr)
library(rlang)
library(rms)
library(pROC)
library(PresenceAbsence)
library(OptimalCutpoints)
library(ggtext)
library(scales)
library(PRROC)
library(patchwork)
```

## Data

```{r}
# Read datasets
data <- read_xlsx("plan/weather_data_final.xlsx")
data_nasa <- read_csv("plan/weather_data_nasa.csv")

# Remove studies 126 to 150
data <- data %>% filter(!study %in% 126:150)
data_nasa <- data_nasa %>% filter(!study %in% 126:150)

# Read predictor
df_predictors <- read_xlsx("plan/df_predictors.xlsx")

location_year <- read.csv("plan/location_year.csv")

location_year$study <- as.character(location_year$study)

df_predictors <- left_join(df_predictors, location_year)
  

```

## Logistic Models

```{r}
# Set up datadist for rms
dd <- datadist(df_predictors)
options(datadist = "dd")

# Convert epidemic to numeric
obs <- as.numeric(as.character(df_predictors$epidemic))

n <- nrow(df_predictors)

```

Fit logistic regression models using predictors of interest. Restricted cubic splines are applied where appropriate to allow for non-linear effects.

#### Logistic model 1 (LM1)

```{r}
# Fit logistic model with restricted cubic splines

m_logistic <- lrm(factor(epidemic) ~ tmin + rcs(rh, 4), 
                  data = df_predictors, x = TRUE, y = TRUE)
```

#### Logistic model 2 (LM2)

```{r}
# Fit logistic model with restricted cubic splines
m_logistic2 <- lrm(factor(epidemic) ~ rcs(rh, 4) + rcs(dew, 3), 
                  data = df_predictors, x = TRUE, y = TRUE)
```

#### Logistic model 3 (LM3)

```{r}
# Fit logistic model with restricted cubic splines

m_logistic3 <- lrm(factor(epidemic) ~ tmin + prec2, 
                  data = df_predictors, x = TRUE, y = TRUE)
```

### LM performance

Evaluate models fit using Cox-Snell and Nagelkerke R², Brier Score, ROC-AUC, optimal classification threshold, accuracy, and confusion matrix.

```{r}
evaluate_logistic <- function(model, data, resp_col, B_boot = 1000) {
  n <- nrow(data)
  actual <- data[[resp_col]]
  predicted_prob <- predict(model, type = "fitted")
  
  # Log-likelihoods
  ll_null <- logLik(glm(as.formula(paste(resp_col, "~ 1")), data = data, family = binomial()))
  ll_full <- logLik(model)
  
  # R²
  cs_r2 <- 1 - exp((2 / n) * (ll_null - ll_full))
  nag_r2 <- cs_r2 / (1 - exp((2 / n) * as.numeric(ll_null)))
  
  # Brier
  brier <- mean((predicted_prob - actual)^2)
  
  # ROC-AUC
  roc_obj <- pROC::roc(actual, predicted_prob)
  auc_val <- pROC::auc(roc_obj)
  
  # Optimal threshold
  preds <- data.frame(1, actual, predicted_prob)
  opt_thresh <- optimal.thresholds(preds)$predicted_prob[3]
  predicted_class <- ifelse(predicted_prob > opt_thresh, 1, 0)
  accuracy <- mean(predicted_class == actual)
  
  # Confusion matrix
  conf <- caret::confusionMatrix(
    factor(predicted_class), 
    factor(actual), 
    mode = "everything", 
    positive = "1"
  )
  
  # PR-AUC via bootstrap
  pr_auc_fun <- function(y, p) {
    y <- as.integer(y)
    if (length(unique(y)) < 2) return(NA_real_)
    PRROC::pr.curve(scores.class0 = p[y == 1],
                    scores.class1 = p[y == 0],
                    curve = FALSE)$auc.integral
  }
  
  pr_apparent <- pr_auc_fun(actual, predicted_prob)
  opt_vec <- numeric(B_boot)
  for (b in 1:B_boot) {
    idx_boot <- sample.int(n, replace = TRUE)
    dat_boot <- data[idx_boot, , drop = FALSE]
    fit_b <- update(model, data = dat_boot)
    y_boot <- dat_boot[[resp_col]]
    p_boot <- predict(fit_b, type = "fitted")
    p_test_orig <- predict(fit_b, newdata = data, type = "fitted")
    opt_vec[b] <- pr_auc_fun(y_boot, p_boot) - pr_auc_fun(actual, p_test_orig)
}
  pr_corrected <- pr_apparent - mean(na.omit(opt_vec))
  
  list(
    cs_r2 = cs_r2,
    nag_r2 = nag_r2,
    brier = brier,
    auc_roc = auc_val,
    accuracy = accuracy,
    conf_matrix = conf,
    pr_auc = pr_corrected,
    opt_threshold = opt_thresh
  )
}
```

#### LM1

```{r}
res_LM1 <- evaluate_logistic(m_logistic, df_predictors, "epidemic")
```

#### LM2

```{r}
res_LM2 <- evaluate_logistic(m_logistic2, df_predictors, "epidemic")
```

#### LM3

```{r}
res_LM3 <- evaluate_logistic(m_logistic3, df_predictors, "epidemic")
```

### Models validation

Perform internal validation using bootstrap and cross-validation. PR-AUC is calculated with bootstrap optimism correction to estimate the expected performance on new data.

#### LM1

```{r}
predicted_prob <- predict(m_logistic, type = "fitted")
actual <- df_predictors$epidemic

# Calibration: bootstrap and cross-validation
cal_boot <- calibrate(m_logistic, method = "boot", B = 1000)
plot(cal_boot, main = "Calibration Plot (Bootstrap)", col = "red")

cal_cv <- calibrate(m_logistic, method = "crossvalidation", B = 10)
plot(cal_cv, main = "Calibration Plot (Cross-validation)")

# rms-style calibration plot
val.prob(predicted_prob, actual, pl = TRUE, smooth = TRUE)

# Internal validation
validation_boot <- validate(m_logistic, method = "boot", B = 1000)
print(validation_boot)

validation_cv <- validate(m_logistic, method = "crossvalidation", B = 10)
print(validation_cv)

```

```{r}
pr_auc_bootstrap <- function(model, data, response, B = 1000, seed = 123) {
  set.seed(seed)
  
  # Internal function to compute PR-AUC
  pr_auc_fun <- function(y, p) {
    y <- as.integer(y)
    if (length(unique(y)) < 2) return(NA_real_)
    PRROC::pr.curve(
      scores.class0 = p[y == 1],
      scores.class1 = p[y == 0],
      curve = FALSE
    )$auc.integral
  }
  
  # Extract response and predicted probabilities
  y_full <- data[[response]]
  p_full <- predict(model, type = "fitted")
  
  # Apparent PR-AUC
  pr_apparent <- pr_auc_fun(y_full, p_full)
  
  # Bootstrap to estimate optimism
  n <- nrow(data)
  opt_vec <- numeric(B)
  for (b in 1:B) {
    idx_boot <- sample.int(n, replace = TRUE)
    dat_boot <- data[idx_boot, , drop = FALSE]
    fit_b <- update(model, data = dat_boot)
    
    y_boot <- dat_boot[[response]]
    p_boot <- predict(fit_b, type = "fitted")
    pr_apparent_b <- pr_auc_fun(y_boot, p_boot)
    
    p_test_on_orig <- predict(fit_b, newdata = data, type = "fitted")
    pr_test_b <- pr_auc_fun(y_full, p_test_on_orig)
    
    opt_vec[b] <- pr_apparent_b - pr_test_b
  }
  
  opt_mean <- mean(na.omit(opt_vec))
  pr_corrected <- pr_apparent - opt_mean
  
  # Bootstrap percentile confidence interval
  pr_test_vec <- numeric(B)
  for (b in 1:B) {
    idx_boot <- sample.int(n, replace = TRUE)
    dat_boot <- data[idx_boot, , drop = FALSE]
    fit_b <- update(model, data = dat_boot)
    p_test_on_orig <- predict(fit_b, newdata = data, type = "fitted")
    pr_test_vec[b] <- pr_auc_fun(y_full, p_test_on_orig)
  }
  
  ci <- quantile(na.omit(pr_test_vec), c(0.025, 0.975))
  
  # Return results as a list
  list(
    pr_apparent = pr_apparent,
    pr_corrected = pr_corrected,
    optimism = opt_mean,
    ci_95 = ci
  )
}

```

```{r}
res_LM1 <- pr_auc_bootstrap(m_logistic, df_predictors, "epidemic")
res_LM1
```

#### LM2

```{r}
predicted_prob <- predict(m_logistic2, type = "fitted")
actual <- df_predictors$epidemic

# Calibration: bootstrap and cross-validation
cal_boot <- calibrate(m_logistic2, method = "boot", B = 1000)
plot(cal_boot, main = "Calibration Plot (Bootstrap)", col = "red")

cal_cv <- calibrate(m_logistic2, method = "crossvalidation", B = 10)
plot(cal_cv, main = "Calibration Plot (Cross-validation)")

# rms-style calibration plot
val.prob(predicted_prob, actual, pl = TRUE, smooth = TRUE)

# Internal validation
validation_boot <- validate(m_logistic2, method = "boot", B = 1000)
print(validation_boot)


validation_cv <- validate(m_logistic2, method = "crossvalidation", B = 10)
print(validation_cv)

```

```{r}
res_LM2 <- pr_auc_bootstrap(m_logistic2, df_predictors, "epidemic")
res_LM2
```

#### LM3

```{r}
predicted_prob <- predict(m_logistic3, type = "fitted")
actual <- df_predictors$epidemic

# Calibration: bootstrap and cross-validation
cal_boot <- calibrate(m_logistic3, method = "boot", B = 1000)
plot(cal_boot, main = "Calibration Plot (Bootstrap)", col = "red")

cal_cv <- calibrate(m_logistic3, method = "crossvalidation", B = 10)
plot(cal_cv, main = "Calibration Plot (Cross-validation)")

# rms-style calibration plot
val.prob(predicted_prob, actual, pl = TRUE, smooth = TRUE)

# Internal validation
validation_boot <- validate(m_logistic3, method = "boot", B = 1000)
print(validation_boot)

validation_cv <- validate(m_logistic3, method = "crossvalidation", B = 10)
print(validation_cv)
```

```{r}
res_LM3 <- pr_auc_bootstrap(m_logistic3, df_predictors, "epidemic")
res_LM3
```

## Ensemble Models

```{r}
# Fit base models and get probabilities
models <- list(
  model1 = lrm(factor(epidemic) ~ tmin + rcs(rh, 4), data = df_predictors, x = TRUE, y = TRUE),
  model2 = lrm(factor(epidemic) ~ rcs(rh, 4) + rcs(dew, 3), data = df_predictors, x = TRUE, y = TRUE),
  model3 = lrm(factor(epidemic) ~ tmin + prec2, data = df_predictors, x = TRUE, y = TRUE)
)

# Predicted probabilities
p1 <- predict(models$model1, type = "fitted")
p2 <- predict(models$model2, type = "fitted")
p3 <- predict(models$model3, type = "fitted")
```

Create ensemble predictions:

#### Unweighted

```{r}
# Simple mean of predicted probabilities

ensemble_unw   <- (p1 + p2 + p3) / 3
```

#### Majorityard vote

```{r}
# Define cutpoints for each base model
cut_p1 <- 0.530
cut_p2 <- 0.51
cut_p3 <- 0.460

# Convert predicted probabilities to binary classification
class_p1 <- ifelse(p1 >= cut_p1, 1, 0)
class_p2 <- ifelse(p2 >= cut_p2, 1, 0)
class_p3 <- ifelse(p3 >= cut_p3, 1, 0)

# Majority vote function
hard_vote <- function(...) {
  votes <- c(...)
  if (sum(votes) >= ceiling(length(votes)/2)) {
    return(1)
  } else {
    return(0)
  }
}

# Apply hard vote across observations
ensemble_hard <- mapply(hard_vote, class_p1, class_p2, class_p3)
```

#### Stacked ensemble

```{r}
# Use base model predictions as features in a meta logistic regression model

stack_data     <- data.frame(p1 = p1, p2 = p2, p3 = p3, epidemic = factor(df_predictors$epidemic))

meta_model     <- glm(epidemic ~ p1 + p2 + p3, data = stack_data, family = binomial)
ensemble_stack <- predict(meta_model, type = "response")
print(meta_model)
```

```{r}
df_preds <- data.frame(
  epidemic = df_predictors$epidemic,  # variável resposta observada
  LM1   = predict(models$model1, type = "fitted"),
  LM2   = predict(models$model2, type = "fitted"),
  LM3   = predict(models$model3, type = "fitted"),
  UNW   = ensemble_unw,
  MJT   = ensemble_hard,
  STACK = predict(meta_model, type = "response")
)

write_xlsx(df_preds, "plan/df_preds.xlsx")
```

### Ensenble models metrics

Metrics: Brier score, optimal threshold, accuracy, confusion matrix, Youden index

```{r}
evaluate_ensemble <- function(predicted_prob, actual) {
  brier_score <- mean((predicted_prob - actual)^2)
  
  # Optimal threshold
  preds <- data.frame(1, actual, predicted_prob)
  o <- optimal.thresholds(preds)
  threshold <- o$predicted_prob[3]
  
  # Binary classification and accuracy
  predicted <- ifelse(predicted_prob > threshold, 1, 0)
  accuracy <- mean(predicted == actual)
  
  # Confusion matrix
  cmax <- confusionMatrix(
    data = as.factor(predicted),
    reference = as.factor(actual),
    mode = "everything",
    positive = "1"
  )
  Sensitivity <- cmax$byClass["Sensitivity"]
  Specificity <- cmax$byClass["Specificity"]
  Youden <- Sensitivity + Specificity - 1
  
  list(
    Brier = brier_score,
    Threshold = threshold,
    Accuracy = accuracy,
    Sensitivity = Sensitivity,
    Specificity = Specificity,
    Youden = Youden
  )
}
```

#### Unweighted

```{r}
actual <- df_predictors$epidemic
res_unw <- evaluate_ensemble(ensemble_unw, actual)
res_unw
```

#### Majority vote

```{r}
res_hard <- evaluate_ensemble(ensemble_hard, actual)
res_hard
```

#### Stacked

```{r}
res_stack <- evaluate_ensemble(ensemble_stack, actual)
res_stack
```

### Ensemble models validation

Bootstrap procedure to estimate variability of AUC, Brier score, and PR-AUC.

Function adapted for ensemble bootstrap metrics (Unweighted / Hard vote)

```{r}
ensemble_metrics_bootstrap <- function(data, type = c("unweighted", "hard"), 
                                       B = 1000, cutpoints = c(0.53, 0.51, 0.46), seed = 123) {
  set.seed(seed)
  type <- match.arg(type)
  
  n <- nrow(data)
  aucs <- numeric(B)
  briers <- numeric(B)
  pr_aucs <- numeric(B)
  
  for (b in 1:B) {
    # 1️⃣ Resample
    idx <- sample(1:n, size = n, replace = TRUE)
    boot_data <- data[idx, ]
    
    # 2️⃣ Fit base models
    m1 <- lrm(factor(epidemic) ~ tmin + rcs(rh, 4), data = boot_data, x=TRUE, y=TRUE)
    m2 <- lrm(factor(epidemic) ~ rcs(rh, 4) + rcs(dew, 3), data = boot_data, x=TRUE, y=TRUE)
    m3 <- lrm(factor(epidemic) ~ tmin + prec2, data = boot_data, x=TRUE, y=TRUE)
    
    # 3️⃣ Predictions
    p1b <- predict(m1, type = "fitted")
    p2b <- predict(m2, type = "fitted")
    p3b <- predict(m3, type = "fitted")
    
    # 4️⃣ Ensemble
    if (type == "unweighted") {
      ens_b <- (p1b + p2b + p3b)/3
    } else if (type == "hard") {
      class_p1 <- ifelse(p1b >= cutpoints[1], 1, 0)
      class_p2 <- ifelse(p2b >= cutpoints[2], 1, 0)
      class_p3 <- ifelse(p3b >= cutpoints[3], 1, 0)
      # Softified probability for metrics (proportion of votes)
      ens_b <- (class_p1 + class_p2 + class_p3)/3
    }
    
    # 5️⃣ Metrics
    aucs[b] <- as.numeric(roc(boot_data$epidemic, ens_b)$auc)
    briers[b] <- mean((ens_b - as.numeric(boot_data$epidemic))^2)
    pr <- pr.curve(scores.class0 = ens_b[boot_data$epidemic==1],
                   scores.class1 = ens_b[boot_data$epidemic==0],
                   curve = FALSE)
    pr_aucs[b] <- pr$auc.integral
  }
  
  list(AUC = aucs, Brier = briers, PR_AUC = pr_aucs)
}
```

#### Unweighted

```{r}
res_unweighted <- ensemble_metrics_bootstrap(df_predictors, type = "unweighted", B = 1000)
mean(res_unweighted$AUC)
mean(res_unweighted$Brier)
mean(res_unweighted$PR_AUC)
```

#### Majority vote

```{r}
res_hard <- ensemble_metrics_bootstrap(df_predictors, type = "hard", B = 1000)
mean(res_hard$AUC)
mean(res_hard$Brier)
mean(res_hard$PR_AUC)
```

#### Stacked

Stacked ensemble calibration and validation

```{r}
dd <- datadist(stack_data)

options(datadist = "dd")

fit_stacked <- lrm(epidemic ~ p1 + p2 + p3, data = stack_data, x = TRUE, y = TRUE)

# Calibration: bootstrap and cross-validation
cal_boot <- calibrate(fit_stacked, method = "boot", B = 1000)
plot(cal_boot, main = "Calibration Plot (Bootstrap)", col = "red")

cal_cv <- calibrate(fit_stacked, method = "crossvalidation", B = 10)
plot(cal_cv, main = "Calibration Plot (Cross-validation)")

# rms-style calibration plot
val.prob(predicted_prob, actual, pl = TRUE, smooth = TRUE)

# Internal validation
validation_boot <- validate(fit_stacked, method = "boot", B = 1000)
print(validation_boot)

validation_cv <- validate(fit_stacked, method = "crossvalidation", B = 10)
print(validation_cv)

# Extract corrected AUC from Dxy
Dxy <- validation_boot["Dxy", "index.corrected"]
B <- validation_boot["B", "index.corrected"]
auc <- (Dxy + 1) / 2
B
```



```{r}
table(df_predictors$year, df_predictors$epidemic)

table(df_predictors$location, df_predictors$epidemic)
```


## LOYO e LOLO cross validation

```{r}

# ======================================================
# ============================================================
# LOYO + LOLO validation for STACKED ensemble (robust)
# - trains base LMs + meta-model within each fold
# - uses Youden threshold from pROC::coords (stable)
# ============================================================

pr_auc_fun <- function(y, p) {
  y <- as.integer(y)
  if (length(unique(y)) < 2) return(NA_real_)
  PRROC::pr.curve(scores.class0 = p[y == 1],
                  scores.class1 = p[y == 0],
                  curve = FALSE)$auc.integral
}

compute_metrics <- function(y, p) {
  y <- as.integer(y)
  p <- as.numeric(p)

  if (length(y) != length(p)) stop("y and p have different lengths.")

  # Brier sempre
  brier <- mean((p - y)^2)

  roc_auc <- pr_auc <- threshold <- sens <- spec <- acc <- NA_real_

  if (length(unique(y)) >= 2) {
    roc_obj <- pROC::roc(y, p, quiet = TRUE)
    roc_auc <- as.numeric(pROC::auc(roc_obj))

    pr_auc <- PRROC::pr.curve(scores.class0 = p[y == 1],
                              scores.class1 = p[y == 0],
                              curve = FALSE)$auc.integral

    # ---- robust threshold by Youden ----
    # Try pROC::coords first, but safely coerce whatever comes back
    thr <- try({
      cc <- pROC::coords(
        roc_obj,
        x = "best",
        best.method = "youden",
        ret = c("threshold", "sensitivity", "specificity"),
        transpose = FALSE
      )

      # coords may return vector, matrix, or data.frame; normalize to data.frame
      if (is.vector(cc) && !is.list(cc)) {
        # named numeric vector
        thr_val <- as.numeric(cc[["threshold"]])
      } else {
        cc_df <- as.data.frame(cc)
        thr_val <- suppressWarnings(as.numeric(cc_df$threshold[1]))
        if (is.na(thr_val)) {
          # sometimes column name differs; take first element
          thr_val <- suppressWarnings(as.numeric(unlist(cc_df)[1]))
        }
      }
      thr_val
    }, silent = TRUE)

    # Fallback if coords fails
    if (inherits(thr, "try-error") || is.na(thr)) {
      # manual Youden: choose threshold that maximizes (sens + spec - 1)
      # use all unique p as candidate thresholds (efficient enough for n=125)
      cand <- sort(unique(p))
      youden <- sapply(cand, function(t) {
        yhat <- ifelse(p > t, 1, 0)
        TP <- sum(yhat == 1 & y == 1)
        TN <- sum(yhat == 0 & y == 0)
        FP <- sum(yhat == 1 & y == 0)
        FN <- sum(yhat == 0 & y == 1)
        sens0 <- if ((TP + FN) > 0) TP/(TP + FN) else NA_real_
        spec0 <- if ((TN + FP) > 0) TN/(TN + FP) else NA_real_
        (sens0 + spec0 - 1)
      })
      threshold <- cand[which.max(youden)]
    } else {
      threshold <- thr
    }

    yhat <- ifelse(p > threshold, 1, 0)

    TP <- sum(yhat == 1 & y == 1)
    TN <- sum(yhat == 0 & y == 0)
    FP <- sum(yhat == 1 & y == 0)
    FN <- sum(yhat == 0 & y == 1)

    sens <- if ((TP + FN) > 0) TP / (TP + FN) else NA_real_
    spec <- if ((TN + FP) > 0) TN / (TN + FP) else NA_real_
    acc  <- (TP + TN) / length(y)
  }

  tibble(
    roc_auc = roc_auc,
    pr_auc  = pr_auc,
    brier   = brier,
    threshold = threshold,
    sensitivity = sens,
    specificity = spec,
    accuracy = acc
  )
}


fit_predict_stacked <- function(train, test) {
  # base models
  m1 <- lrm(factor(epidemic) ~ tmin + rcs(rh, 4), data = train, x=TRUE, y=TRUE)
  m2 <- lrm(factor(epidemic) ~ rcs(rh, 4) + rcs(dew, 3), data = train, x=TRUE, y=TRUE)
  m3 <- lrm(factor(epidemic) ~ tmin + prec2, data = train, x=TRUE, y=TRUE)

  # train preds for meta-model
  p1_tr <- as.numeric(predict(m1, type="fitted"))
  p2_tr <- as.numeric(predict(m2, type="fitted"))
  p3_tr <- as.numeric(predict(m3, type="fitted"))

  meta <- glm(factor(train$epidemic) ~ p1_tr + p2_tr + p3_tr, family = binomial)

  # test preds for meta-model
  p1_te <- as.numeric(predict(m1, newdata=test, type="fitted"))
  p2_te <- as.numeric(predict(m2, newdata=test, type="fitted"))
  p3_te <- as.numeric(predict(m3, newdata=test, type="fitted"))

  # ensure lengths match
  if (!all(c(length(p1_te), length(p2_te), length(p3_te)) == nrow(test))) {
    stop("Prediction length mismatch: check factor levels or data issues in test set.")
  }

  p_stack <- as.numeric(predict(meta,
                               newdata = data.frame(p1_tr=p1_te, p2_tr=p2_te, p3_tr=p3_te),
                               type = "response"))
  p_stack
}

run_group_cv <- function(data, group_col = c("year","location")) {
  group_col <- match.arg(group_col)

  dat <- data %>%
    mutate(
      epidemic = as.integer(as.character(epidemic)),
      .group = as.character(.data[[group_col]])
    ) %>%
    filter(!is.na(.group))

  groups <- sort(unique(dat$.group))

  map_dfr(groups, function(g) {
    train <- dat %>% filter(.group != g)
    test  <- dat %>% filter(.group == g)

    # If train lacks both classes, cannot fit meaningful classifier
    if (length(unique(train$epidemic)) < 2) {
      return(tibble(group=g, n_test=nrow(test), status="skip: train single class"))
    }

    # Predict probabilities on test
    p <- fit_predict_stacked(train, test)

    # If test is single-class, AUC undefined, but Brier still meaningful
    if (length(unique(test$epidemic)) < 2) {
      brier <- mean((as.numeric(p) - as.integer(test$epidemic))^2)
      return(tibble(group=g, n_test=nrow(test), status="ok (single-class test)",
                    roc_auc=NA_real_, pr_auc=NA_real_, brier=brier,
                    threshold=NA_real_, sensitivity=NA_real_, specificity=NA_real_, accuracy=NA_real_))
    }

    met <- compute_metrics(test$epidemic, p)
    bind_cols(tibble(group=g, n_test=nrow(test), status="ok"), met)
  })
}

summarize_cv <- function(res) {
  ok <- res %>% filter(grepl("^ok", status))
  tibble(
    folds_total = nrow(res),
    folds_ok = nrow(ok),
    n_test_total = sum(ok$n_test, na.rm=TRUE),
    roc_auc_mean = mean(ok$roc_auc, na.rm=TRUE),
    roc_auc_median = median(ok$roc_auc, na.rm=TRUE),
    roc_auc_IQR = IQR(ok$roc_auc, na.rm=TRUE),
    pr_auc_mean = mean(ok$pr_auc, na.rm=TRUE),
    pr_auc_median = median(ok$pr_auc, na.rm=TRUE),
    pr_auc_IQR = IQR(ok$pr_auc, na.rm=TRUE),
    brier_mean = mean(ok$brier, na.rm=TRUE),
    brier_median = median(ok$brier, na.rm=TRUE),
    brier_IQR = IQR(ok$brier, na.rm=TRUE),
    sens_mean = mean(ok$sensitivity, na.rm=TRUE),
    spec_mean = mean(ok$specificity, na.rm=TRUE),
    acc_mean  = mean(ok$accuracy, na.rm=TRUE)
  )
}

# ---- run ----
df_predictors <- df_predictors %>%
  mutate(year = as.character(year),
         location = as.character(location))

cv_loyo <- run_group_cv(df_predictors, "year")
cv_lolo <- run_group_cv(df_predictors, "location")

sum_loyo <- summarize_cv(cv_loyo)
sum_lolo <- summarize_cv(cv_lolo)

print("==== LOYO (STACK) ===="); print(sum_loyo)
print("==== LOLO (STACK) ===="); print(sum_lolo)

write_xlsx(list(LOYO=cv_loyo, LOLO=cv_lolo,
                Summary_LOYO=sum_loyo, Summary_LOLO=sum_lolo),
           "plan/cv_STACK_LOYO_LOLO.xlsx")

cv_loyo %>% count(status)
cv_lolo %>% count(status)

```


