---
title: ""
message: false
warning: false
---

# Weather Data Script

This repository is intended to demonstrate the **structure and workflow of the code** used in the project.\
To keep the repository lightweight and accessible on GitHub, the original input data files (provided by Xavier et al. 2022) have been **excluded** because they exceed GitHub's file size limits.

The corresponding lines in the code that load these large `.nc` files are commented out (using `#`). This way, the code can still be inspected and followed without requiring the full dataset. However, the scripts will not reproduce the complete analysis unless the data are available locally.

## Libraries

Load Required Libraries

```{r}
library(dplyr)
library(purrr)
library(gsheet)
library(raster)
library(ncdf4)
library(lubridate)
library(readxl)
library(writexl)
library(caret)
library(tidyr)
library(r4pde)
library(refund)
library(ggplot2)
library(readr)
```

## Experimental Metadata

Load Experimental Metadata

```{r}
# Load trial metadata from Google Sheets
gib <- gsheet2tbl("https://docs.google.com/spreadsheets/d/1MBiKsosQ8Hob6LkS65_1pPU25hx1CO9i42Sm_xf28ww/edit?gid=0#gid=0")
```

Load experimental metadata and convert dates to Date format. Create a 57-day window (±28 days) around flowering for extracting climate data.

```{r}
# Convert dates and ensure study is a factor
gib2 <- gib |> 
  mutate(
    planting_date = as.Date(planting_date, format = "%Y-%m-%d"),
    flowering = as.Date(flowering, format = "%Y-%m-%d"),
    study = as.factor(study)
  )

# Define ±28-day window around flowering for climate data
gib3 <- gib2 |>  mutate(pd28 = flowering + 28,
                        pd00 = flowering - 28,
                        year = as.numeric(year))

```

## Climate Data Extraction from NetCDF (Xavier)

The full dataset can be obtained from Xavier et al. (2022) at the following link: [Xavier Dataset – Download here](https://sites.google.com/site/alexandrecandidoxavierufes/brazilian-daily-weather-gridded-data)

After downloading, place the files in the `data/` directory of this repository and uncomment the relevant lines in the scripts to run the full analysis.

Function to extract climate data from NetCDF files:

```{r}
## Function to extract daily climate variables from NetCDF for specified trials

#extract_nc_var <- function(netcdf_path, gib, varname, outfile) {
#  # Open NetCDF file
#  nc_data <- nc_open(netcdf_path)
#  
#  # Convert NetCDF time to Date
#  time_var   <- ncvar_get(nc_data, "time")
#  time_units <- ncatt_get(nc_data, "time", "units")$value
#  origin_date <- as.POSIXct(sub("hours since ", "", time_units), tz = "UTC")
#  nc_dates   <- as.Date(origin_date + as.difftime(time_var, units = "hours"))
  
#  # Extract grid coordinates
#  lon_vals <- ncvar_get(nc_data, "longitude")
#  lat_vals <- ncvar_get(nc_data, "latitude")
#  
#  # Internal function to extract data for each row of 'gib'
#  get_data_for_row <- function(row, line_number) {
#    start_date <- row$pd00
#    end_date   <- row$pd28
#    date_seq   <- seq.Date(start_date, end_date, by = "day")
#    date_idx   <- match(date_seq, nc_dates)
#    date_idx   <- date_idx[!is.na(date_idx)]
    
#    if (length(date_idx) == 0) return(NULL)
    
#    # Find closest grid point
#    lon_idx <- which.min(abs(lon_vals - row$lon))
#    lat_idx <- which.min(abs(lat_vals - row$lat))
    
#    # Extract values
#    values <- sapply(date_idx, function(i) {
#      ncvar_get(nc_data, varname,
#                start = c(lon_idx, lat_idx, i),
#                count = c(1, 1, 1))
#    })
    
#    if (length(values) == 0 || all(is.na(values))) return(NULL)
    
#    # Build dataframe
#    df <- data.frame(
#      date          = date_seq[!is.na(match(date_seq, nc_dates))],
#      value         = values,
#      study         = row$study,
#      location      = row$location,
#      state         = row$state,
#      year          = format(start_date, "%Y"),
#      planting_date = row$planting_date
#    )
    
#    names(df)[2] <- varname
#    return(df)
#  }
  
#  # Apply function for each row of 'gib'
#  result <- gib %>%
#    split(1:nrow(.)) %>%
#    map2_dfr(1:nrow(gib), ~ get_data_for_row(.x, .y), .id = "id")
  
#  # Save as Excel
#  write_xlsx(result, outfile)
  
#  # Close NetCDF
#  nc_close(nc_data)
#  return(result)
#}

```

This function extracts a specific climate variable (e.g., precipitation, Tmax, Tmin, RH) from a NetCDF file for each trial, using the nearest grid point to the trial location. Results are saved as Excel.

Now, extract climate variables from Xavier dataset for old and new periods:

```{r}
#   # Extracting Climate Variables

#   # Precipitation
#prec_old <- extract_nc_var("data/pr_19810101_20001231_BR-DWGD_UFES_UTEXAS_v_3.2.3.nc", gib3, "pr", "data/br_dwgd_prec_old.xlsx")

#prec_new <- extract_nc_var("data/pr_20010101_20240320_BR-DWGD_UFES_UTEXAS_v_3.2.3.nc", gib3, "pr", "data/br_dwgd_prec.xlsx")

#   # Tmax
#tmax_old <- extract_nc_var("data/Tmax_19810101_20001231_BR-DWGD_UFES_UTEXAS_v_3.2.3.nc", gib3, "Tmax", "data/br_dwgd_tmax_old.xlsx")

#tmax_new <- extract_nc_var("data/Tmax_20010101_20240320_BR-DWGD_UFES_UTEXAS_v_3.2.3.nc", gib3, "Tmax", "data/br_dwgd_tmax.xlsx")

#   # Tmin
#tmin_old <- extract_nc_var("data/Tmin_19810101_20001231_BR-DWGD_UFES_UTEXAS_v_3.2.3.nc", gib3, "Tmin", "data/br_dwgd_tmin_old.xlsx")

#tmin_new <- extract_nc_var("data/Tmin_20010101_20240320_BR-DWGD_UFES_UTEXAS_v_3.2.3.nc", gib3, "Tmin", "data/br_dwgd_tmin.xlsx")

#   # Relative humidity
#rh_old <- extract_nc_var("data/RH_19810101_20001231_BR-DWGD_UFES_UTEXAS_v_3.2.3.nc", gib3, "RH", "data/br_dwgd_rh_old.xlsx")

#rh_new <- extract_nc_var("data/RH_20010101_20240320_BR-DWGD_UFES_UTEXAS_v_3.2.3.nc", gib3, "RH", "data/br_dwgd_rh.xlsx")
```

### Combine Xavier Climate Data

To facilitate the joining process, precipitation, temperature (max and min), and relative humidity (RH) data from both old and new sources were combined externally into "final" Excel files. These files are then read into R.

```{r}
# Read final climate datasets

prec <- read_xlsx("plan/br_dwgd_prec_final.xlsx")
tmax <- read_xlsx("plan/br_dwgd_tmax_final.xlsx")
tmin <- read_xlsx("plan/br_dwgd_tmin _final.xlsx")
rh   <- read_xlsx("plan/br_dwgd_rh_final.xlsx")

# Combine into a single dataframe
weather_data <- prec |> 
  mutate(
    RH2M        = rh$rh,
    T2M_MAX     = tmax$tmax,
    T2M_MIN     = tmin$tmin,
    PRECTOTCORR = prec$precipitation
  ) |> 
  dplyr::select(-study) |> 
  rename(study = id) |> 
  mutate(
    date = as.Date(date),
    study = as.factor(study),
    planting_date = as.Date(planting_date)
  )

# Join with trial metadata
weather_data2 <- left_join(weather_data, gib3, by = "study") |> 
  dplyr::select(
    study, date, location.x, state.x, year.x, lat, lon, planting_date.x, index, flowering, RH2M, T2M_MIN, T2M_MAX, PRECTOTCORR
  )|> 
  rename(location = location.x,
         state = state.x,
         year = year.x,,
         planting_date =  planting_date.x)|> 
  mutate(
    T2M = (T2M_MAX + T2M_MIN) / 2,                 
    DIFF_TMAX_TMIN = T2M_MAX - T2M_MIN,             
    year = as.numeric(year))


write_xlsx(weather_data2, "plan/weather_data2_final.xlsx")

```

Combine precipitation, temperature, and RH into a single dataframe and join with experimental metadata. Compute daily mean temperature and range.

## NASA Power Supplementation (2024 Trials)

Some trials from 2024 did not have climate data available in the Xavier dataset. These trials were supplemented using the NASA Power API.

```{r}
# Select trials missing Xavier climate data
gib4 <- gib3 |> 
  filter(study %in% c(92, 93, 94, 95, 96, 97, 98, 99, 100, 101, 126, 127, 128, 129, 130, 131, 132, 149, 150))

# Prepare locations
trials <- gib4 |> 
  dplyr::select(study, flowering, lon, lat) |> 
  rename(
    longitude = lon,
    latitude = lat
  )

# Retrieve weather from NASA Power
weather_nasa <- get_nasapower(
  data = trials,
  days_around = 28,
  date_col = "flowering",
  pars = c("GWETROOT", "GWETTOP", "T2M", "T2M_MAX", "T2M_MIN", "T2M_RANGE", "RH2M", "PRECTOTCORR", "T2MDEW","WS2M", "PS", "ALLSKY_SFC_SW_DWN", "CLRSKY_SFC_SW_DWN"))

# Save
write_xlsx(weather_nasa, "plan/weather_nasa.xlsx")
write_csv(weather_nasa, "plan/weather_nasa.csv")

# Combine with experimental metadata
trials_weather <- full_join(gib3, weather_nasa)

write_xlsx(trials_weather, "plan/trials_weather.xlsx")
write_csv(trials_weather, "plan/trials_weather.csv")
```

## Final Climate Dataset (Xavier + NASA)

Again to facilitate the joining process, data from Xavier and Nasa Power were combined externally into "dados_clima_completo" Excel files. These files are then read into R.

```{r}
dados_clima <- read_xlsx("plan/dados_clima_completo.xlsx") |> 
  mutate(
    date = as.Date(date),
    flowering = as.Date(flowering)
  )


weather_data_final <- dados_clima |> 
  group_by(study) |> 
  # Calculate the number of days relative to flowering
  mutate(days = as.numeric(date - flowering)) |> 
  ungroup() |> 
  # Classify epidemic vs non-epidemic trials
  mutate(epidemic = if_else(index >= 10, 1, 0),
         # Compute mean temperature and daily temperature range
         T2M = (T2M_MAX + T2M_MIN)/2,
         T2M_RANGE = T2M_MAX - T2M_MIN,
         e_s = 0.6108 * exp((17.27 * T2M) / (T2M + 237.3)),
         e_a = e_s * (RH2M / 100),
         VPD = e_s - e_a)


write_xlsx(weather_data_final, "plan/weather_data_final.xlsx")
write_csv(weather_data_final, "plan/weather_data_final.csv")
```

Compute derived variables such as mean temperature, daily temperature range, vapor pressure (saturation and actual), and VPD for further analysis.

## Additional Nasa Power weather variables

```{r}
# Prepare unique trial locations for NASA Power retrieval

 data1 <- weather_data_final |> 
   group_by(study) |> 
   slice(1) |> 
   dplyr::select(study, flowering, lat, lon) |> 
   rename(latitude = lat, longitude = lon)

# Download additional weather variables from NASA Power
 weather_data <- r4pde::get_nasapower(
    data = data1,
    days_around = 28,
     date_col = "flowering",
    pars = c("GWETROOT", "GWETTOP", "T2M", "T2M_MAX", "T2M_MIN", "T2M_RANGE", "RH2M", "PRECTOTCORR", "T2MDEW","WS2M", "PS", "ALLSKY_SFC_SW_DWN", "CLRSKY_SFC_SW_DWN"))
 
 trials <- data1 |> 
   dplyr::select(study, flowering)
 
 data_nasa <- left_join(trials, weather_data)
 
 epidemic <- weather_data_final |> 
   group_by(study) |> 
   slice(1) |> 
   dplyr::select(epidemic)
 
 
 data_nasa <- data_nasa |> 
  mutate(days = as.numeric(YYYYMMDD - as.Date(flowering)),
         study = as.factor(study)) |> 
   left_join(epidemic) |> 
   mutate(TDD = T2M_MIN - T2MDEW)

 # Save final NASA Power dataset
write_csv(data_nasa, "plan/weather_data_nasa.csv")

```

## ERA5 Reanalysis Supplementation (Studies 1-125)

This section fetches daily climate data from the ERA5 reanalysis via the Open-Meteo API for studies 1 to 125, using a ±28-day window around the flowering date.

```{r}
# Filter studies 1 to 125
gib_era5 <- gib3 |> 
  filter(as.numeric(as.character(study)) %in% 1:125)

# Function to fetch ERA5 data from Open-Meteo API
get_open_meteo_era5 <- function(row) {
  lat <- row$lat
  lon <- row$lon
  start_date <- row$pd00
  end_date <- row$pd28
  study_id <- row$study
  
  # Add a delay to avoid rate limiting
  Sys.sleep(2)
  
  # Construct API URL for essential variables only
  base_url <- "https://archive-api.open-meteo.com/v1/archive"
  query <- paste0(
    "?latitude=", lat,
    "&longitude=", lon,
    "&start_date=", start_date,
    "&end_date=", end_date,
    "&daily=temperature_2m_max,temperature_2m_min,temperature_2m_mean,precipitation_sum,relative_humidity_2m_mean,dew_point_2m_mean",
    "&timezone=GMT&format=csv"
  )
  
  full_url <- paste0(base_url, query)
  
  # Debug print
  message(paste("Fetching Study", study_id, "from:", full_url))
  
  # Multi-strategy download function: tries wininet -> curl pkg -> system curl
  fetch_url_robust <- function(url, dest) {
    
    # Strategy 1: wininet (Windows native, same as browser)
    tryCatch({
      download.file(url, dest, method = "wininet", mode = "wb", quiet = TRUE)
      if(file.exists(dest) && file.info(dest)$size > 0) {
        message("  -> Downloaded using wininet")
        return(TRUE)
      }
    }, error = function(e) { })
    
    # Strategy 2: curl package with IPv4 + SSL bypass
    tryCatch({
      h <- new_handle()
      handle_setopt(h, ssl_verifypeer = 0L, ssl_verifyhost = 0L, ipresolve = 1L)
      curl_download(url, dest, handle = h, quiet = TRUE)
      if(file.exists(dest) && file.info(dest)$size > 0) {
        message("  -> Downloaded using curl package (IPv4)")
        return(TRUE)
      }
    }, error = function(e) { })
    
    # Strategy 3: System curl executable
    tryCatch({
      exit_code <- system2("curl", c("-L", "-k", "-s", "-o", shQuote(dest), shQuote(url)), stdout = FALSE, stderr = FALSE)
      if(exit_code == 0 && file.exists(dest) && file.info(dest)$size > 0) {
        message("  -> Downloaded using system curl")
        return(TRUE)
      }
    }, error = function(e) { })
    
    return(FALSE)
  }
  
  # Main fetch logic
  tryCatch({
    
    base_url <- "https://archive-api.open-meteo.com/v1/archive"
    
    # Construct URL manually
    query_parts <- c(
      paste0("latitude=", lat),
      paste0("longitude=", lon),
      paste0("start_date=", start_date), 
      paste0("end_date=", end_date),
      paste0("daily=", URLencode("temperature_2m_max,temperature_2m_min,temperature_2m_mean,precipitation_sum,relative_humidity_2m_mean,dew_point_2m_mean")),
      "timezone=GMT",
      "format=csv"
    )
    
    full_url <- paste0(base_url, "?", paste(query_parts, collapse = "&"))
    
    temp_file <- tempfile(fileext = ".csv")
    
    # Try all strategies
    if(!fetch_url_robust(full_url, temp_file)) {
      stop("All download strategies failed")
    }

    # Read the downloaded CSV using robust column indexing
    df <- read_csv(temp_file, skip = 3, show_col_types = FALSE) |> 
      mutate(study = study_id,
             flowering = row$flowering) |> 
      dplyr::select(
        study,
        flowering,
        date = 1,        # time
        T2M_MAX = 2,     # temperature_2m_max
        T2M_MIN = 3,     # temperature_2m_min
        T2M = 4,         # temperature_2m_mean
        PRECTOTCORR = 5, # precipitation_sum
        RH2M = 6,        # relative_humidity_2m_mean
        T2MDEW = 7       # dew_point_2m_mean
      ) |> 
      mutate(
        date = as.Date(date),
        T2M_RANGE = T2M_MAX - T2M_MIN
      )
    
    # Clean up
    unlink(temp_file)
    
    return(df)
  }, error = function(e) {
    message(paste("Error fetching Study", study_id, "::", e$message))
    return(NULL)
  })
}

# Iterate through studies 1 to 125
weather_data_era5 <- gib_era5 |> 
  split(1:nrow(gib_era5)) |> 
  map_dfr(get_open_meteo_era5)

era5 <- weather_data_era5 |> 
 mutate(days = as.numeric(date - flowering)) 




# Save the results
write_xlsx(weather_data_era5, "plan/weather_data_era5.xlsx")
write_csv(weather_data_era5, "plan/weather_data_era5.csv")



final <- read_csv("plan/weather_data_final.csv")
final$study <- as.numeric(final$study)



era5 <- read_csv("plan/weather_data_era5.csv")

final <- final |> 
  filter(study < 126)

era5 <- era5 |> 
  mutate(days = as.numeric(date - flowering))
era5$epidemic <- final$epidemic

final2 <- final |> 
  dplyr::select(study, flowering, date, T2M_MAX, T2M_MIN, T2M, PRECTOTCORR, RH2M, T2M_RANGE, days, epidemic)

era5

final2$T2MDEW <- era5$T2MDEW

final3 <- final2 |> 
  filter(study < 100)

era53 <- era5 |> 
  filter(study > 99)

final_final <- rbind(final3, era53)

write_csv(final_final, "plan/weather_data_final_era5.csv")

```


## Simulation of peak flowering


```{r}

library(dplyr)

weibull_cdf <- function(x, a, b) 1 - exp(-a * x^b)

a_of_T <- function(T) 0.2551 - 0.029*T + 0.0009*T^2
b_of_T <- function(T) -5.7736 + 0.9663*T - 0.0278*T^2

simulate_peak_by_study <- function(df_one,
                                   wheat_type = c("br","winter"),
                                   heading_offset_days = 5,  # heading = flowering - 5
                                   lag_days = 0,             # set 0 if flowering ~ start of extrusion
                                   max_days_after_flowering = 28) {

  wheat_type <- match.arg(wheat_type)

  if (wheat_type == "br") { aE <- 0.0127; bE <- 2.4352 }
  else { aE <- 0.0447; bE <- 2.5041 }

  df_one <- df_one %>%
    arrange(date) %>%
    filter(days >= -heading_offset_days,
           days <= max_days_after_flowering) %>%
    mutate(Tmean = if ("T2M" %in% names(.)) T2M else (T2M_MIN + T2M_MAX)/2)

  flowering0 <- unique(df_one$flowering)[1]
  heading_start <- flowering0 - heading_offset_days

  # axis in days since HEADING (not since flowering)
  df_one <- df_one %>%
    mutate(days_since_heading = as.integer(date - heading_start))

  # 1) heading progress
  df_one <- df_one %>%
    mutate(EspRel_cum   = weibull_cdf(pmax(days_since_heading, 0), aE, bE),
           EspRel_daily = c(EspRel_cum[1], diff(EspRel_cum))) %>%
    mutate(EspRel_daily = pmax(EspRel_daily, 0))

  # 2) anther extrusion: cohorts start after lag_days (relative to heading axis)
  n <- nrow(df_one)
  extruded_daily <- numeric(n)

  for (i in seq_len(n)) {
    cohort <- df_one$EspRel_daily[i]
    if (cohort <= 0) next

    start_j <- i + lag_days
    if (start_j > n) next

    for (j in start_j:n) {
      x <- (j - start_j) + 1L
      Tj <- df_one$Tmean[j]
      aj <- a_of_T(Tj)
      bj <- b_of_T(Tj)

      Fx  <- weibull_cdf(x,  aj, bj)
      Fx1 <- weibull_cdf(x-1, aj, bj)

      extruded_daily[j] <- extruded_daily[j] + cohort * max(0, Fx - Fx1)
    }
  }

  df_one$anther_extruded_daily <- extruded_daily

  # peak AFTER flowering date (optional but usually what you want)
  after_flw <- which(df_one$date >= flowering0)
  peak_i <- after_flw[which.max(df_one$anther_extruded_daily[after_flw])]

  tibble(
    flowering0 = flowering0,
    heading_start = heading_start,
    lag_days = lag_days,
    peak_date = df_one$date[peak_i],
    peak_days_after_flowering = as.numeric(df_one$date[peak_i] - flowering0),
    peak_value = df_one$anther_extruded_daily[peak_i]
  )
}



res_lag0 <- weather_data_final %>%
  group_by(study) %>%
  group_modify(~ simulate_peak_by_study(.x, wheat_type="br",
                                        heading_offset_days=5, lag_days=0)) %>%
  ungroup() %>% mutate(scenario = "heading=flw-5, lag=0")

res_lag5 <- weather_data_final %>%
  group_by(study) %>%
  group_modify(~ simulate_peak_by_study(.x, wheat_type="br",
                                        heading_offset_days=5, lag_days=5)) %>%
  ungroup() %>% mutate(scenario = "heading=flw-5, lag=5")

bind_rows(res_lag0, res_lag5) %>%
  group_by(scenario) %>%
  summarise(mean_peak = mean(peak_days_after_flowering),
            sd_peak   = sd(peak_days_after_flowering),
            .groups="drop")



```

## Flowering curve

```{r}

library(dplyr)
library(ggplot2)

weibull_cdf <- function(x, a, b) 1 - exp(-a * x^b)

# Del Ponte et al. (2004): extrusão como função de T
a_of_T <- function(T) 0.2551 - 0.029*T + 0.0009*T^2
b_of_T <- function(T) -5.7736 + 0.9663*T - 0.0278*T^2

simulate_anthers_series <- function(df_one,
                                    wheat_type = c("br","winter"),
                                    lag_heading_to_extrusion = 5, # artigo
                                    heading_offset_from_flowering = 5, # heading = flowering - 5
                                    anther_lifespan_days = 3,      # sua regra
                                    max_days_after_heading = 25) {

  wheat_type <- match.arg(wheat_type)

  # Espigamento (Weibull) — trigo BR e trigo inverno (artigo)
  if (wheat_type == "br") { aE <- 0.0127; bE <- 2.4352 } else { aE <- 0.0447; bE <- 2.5041 }

  df_one <- df_one %>%
    arrange(date) %>%
    mutate(Tmean = if ("T2M" %in% names(.)) T2M else (T2M_MIN + T2M_MAX)/2)

  flowering0 <- unique(df_one$flowering)[1]
  heading_start <- flowering0 - heading_offset_from_flowering

  # eixo: dias desde o início do espigamento
  df_one <- df_one %>%
    mutate(days_since_heading = as.integer(date - heading_start)) %>%
    filter(days_since_heading >= 0, days_since_heading <= max_days_after_heading) %>%
    arrange(days_since_heading)

  # 1) Espigamento: cumulativo e diário
  df_one <- df_one %>%
    mutate(EspRel_cum   = weibull_cdf(days_since_heading, aE, bE),
           EspRel_daily = c(EspRel_cum[1], diff(EspRel_cum))) %>%
    mutate(EspRel_daily = pmax(EspRel_daily, 0))

  # 2) Extrusão diária agregada por coortes de espigas
  n <- nrow(df_one)
  extruded_daily <- numeric(n)

  for (i in seq_len(n)) {
    cohort <- df_one$EspRel_daily[i]
    if (cohort <= 0) next

    start_j <- i + lag_heading_to_extrusion
    if (start_j > n) next

    for (j in start_j:n) {
      x <- (j - start_j) + 1L  # dias desde início da extrusão desse coorte

      Tj <- df_one$Tmean[j]
      aj <- a_of_T(Tj)
      bj <- b_of_T(Tj)

      Fx  <- weibull_cdf(x,  aj, bj)
      Fx1 <- weibull_cdf(x-1, aj, bj)

      extruded_daily[j] <- extruded_daily[j] + cohort * max(0, Fx - Fx1)
    }
  }

  df_one$anthers_extruded <- extruded_daily

  # normalizar para % (0–100) dentro do estudo (garante soma=100 no período)
  tot_extr <- sum(df_one$anthers_extruded, na.rm = TRUE)
  if (tot_extr > 0) df_one$anthers_extruded_pct <- 100 * df_one$anthers_extruded / tot_extr
  else df_one$anthers_extruded_pct <- 0

  # 3) Anteras PRESENTES: janela móvel de 3 dias após extrusão (sobe e desce)
  # presente no dia t = soma das extrusões em t, t-1, t-2 (lifespan=3)
  present <- numeric(n)
  L <- anther_lifespan_days
  for (t in seq_len(n)) {
    idx <- max(1, t - (L - 1)) : t
    present[t] <- sum(df_one$anthers_extruded_pct[idx], na.rm = TRUE)
  }
  df_one$anthers_present_pct <- present

df_one %>%
  transmute(
    date,
    flowering0 = flowering0,
    heading_start = heading_start,
    days_since_heading,
    days_since_flowering = as.integer(date - flowering0),
    Tmean,
    anthers_extruded_pct,
    anthers_present_pct
  )

}


curves <- weather_data_final %>%
  group_by(study) %>%
  group_modify(~ simulate_anthers_series(.x, wheat_type = "br")) %>%
  ungroup()





mean_curve <- curves %>%
  group_by(days_since_flowering) %>%
  summarise(mean_present = mean(anthers_present_pct, na.rm = TRUE),
            .groups = "drop")

peak_day <- mean_curve$days_since_flowering[
  which.max(mean_curve$mean_present)
]



ggplot() +
  geom_line(data = curves,
            aes(x = days_since_flowering,
                y = anthers_present_pct,
                group = study),
            alpha = 0.20) +
  geom_line(data = mean_curve,
            aes(x = days_since_flowering,
                y = mean_present),
            linewidth = 1.2,
            inherit.aes = FALSE) +
  geom_vline(xintercept = peak_day, linetype = "dashed") +
  labs(x = "Dias após o início do florescimento",
       y = "% de anteras presentes (janela de 3 dias)",
       title = "Dinâmica do florescimento (anteras presentes)") +
  theme_classic()



mean_flux <- curves %>%
  group_by(days_since_flowering) %>%
  summarise(mean_extruded = mean(anthers_extruded_pct, na.rm = TRUE),
            .groups = "drop")

ggplot() +
  geom_line(data = curves,
            aes(x = days_since_flowering,
                y = anthers_extruded_pct,
                group = study),
            alpha = 0.20) +
  geom_line(data = mean_flux,
            aes(x = days_since_flowering,
                y = mean_extruded),
            linewidth = 1.2,
            inherit.aes = FALSE) +
  labs(x = "Dias após o início do florescimento",
       y = "% de anteras extrusadas no dia",
       title = "Extrusão diária de anteras") +
  theme_classic()




mean_curve <- curves %>%
  group_by(days_since_flowering) %>%
  summarise(mean_present = mean(anthers_present_pct, na.rm = TRUE),
            .groups = "drop")

p_curves <- ggplot() +
  geom_line(data = curves,
            aes(x = days_since_flowering,
                y = anthers_present_pct,
                group = study),
            alpha = 0.18,
            color = "grey80") +
  geom_line(data = mean_curve,
            aes(x = days_since_flowering,
                y = mean_present),
            color = "darkorange",
            linewidth = 1.3,
            inherit.aes = FALSE) +
  coord_cartesian(xlim = c(0, 20)) +
  labs(x = "Days after flowering",
       y = "% present anthers") +
  theme_classic()

p_curves


peak_by_study <- curves %>%
  group_by(study) %>%
  slice_max(anthers_present_pct, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  dplyr::select(study, days_since_flowering, anthers_present_pct) %>%
  rename(peak_days_after_flowering = days_since_flowering,
         peak_value = anthers_present_pct)




m  <- mean(peak_by_study$peak_days_after_flowering)
md <- median(peak_by_study$peak_days_after_flowering)
s  <- sd(peak_by_study$peak_days_after_flowering)

p_histo <- ggplot(peak_by_study,
       aes(x = peak_days_after_flowering)) +
  geom_histogram(binwidth = 1,
                 boundary = 0,
                 closed = "left",
                 color = "black",
                 fill = "lightgray") +
  geom_vline(xintercept = m,
             linetype = "dashed",
             linewidth = 1) +
  geom_vline(xintercept = md,
             linetype = "solid",
             linewidth = 1) +
  labs(x = "Days to reach peak flowering",
       y = "Number of studies")+
  theme_classic()


library(patchwork)
(p_curves + p_histo) + plot_annotation(tag_levels = "A")



p_inset <- p_curves +
  inset_element(
    p_histo +
      labs(title = NULL, x = NULL, y = NULL) +
      theme_classic(base_size = 9),
    left = 0.62, bottom = 0.55, right = 0.98, top = 0.98
  )

```